---
title: "submittedPortfolio"
author: "NareshVemula"
date: "November 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

We just used four packages in financial market forecasting, our model is build from base r functions such as random numbers and arithematic operations 

we also optimized the iterations in the simulated model to linear level of n, which reduced the computation required to power of n

```{r}
require(dplyr)
require(TTR)
require(reshape2)
require(plotly)
```

reading the all ticker symbols(stock option names and valuations from the yahoo finance)
```{r, message=FALSE, warning=FALSE}
SYMs <- TTR::stockSymbols()
head(SYMs)
```

This is the function where our final model is built on and the steps involved in our forecasting 
are mentioned below, this function has three parameters stockURL,forecastDate,simulations,render

1. stockURL( stock url consists of the url to getdata from yahoo finance , this has stock option 
symbol, from date for forecasting, to date to be considered)
by default we are considering past one year data for predicting the future values

2. forecastDate( this will be given by the user, this is the date for which he his aiming to 
buy a call option, based on this we forecast the respective stock to check if the call option  
is valuable or not)

3.  simulations (number of simulations required for computing each day's forecasts from random walk modell
by default we are using 1000 simulations)

4.  render is a toggle button for functions defaultStocks and callingRenderingStockForecast

```{r}
stockForecast=function(stockURL,forecastDate,simulations=1000,render=FALSE)
{
#concatenating the url string to a complete url and hitting yahoo finance to get the data
URL <- paste("http://ichart.finance.yahoo.com/table.csv?s=",stockURL,sep ='')
#code for debugging the code, printing the url
print(URL)
#reading the stock options from the yahoo 
stock <- read.csv(URL)
#converting stock date from yahoo finance into date format
stock$Date <- as.Date(stock$Date, "%Y-%m-%d")
#monte carlo simulations and modeling
#getting the current stock price
currentStockValue=stock$Close[1]
#taking one day lag values into another vector to compute the increase percentage of stock 
laggedStockValue=stock$Close[2:nrow(stock)]
#computing the increase percentage of stock from yesterday to today
percentIncrease=stock$Close[1:(nrow(stock)-1)]/laggedStockValue
#mean increase in the stock price from past one year
meanIncrease=mean(percentIncrease)
#standard deviation of the stock price from past one year
stdIncrease=sd(percentIncrease)
#number of time periods we need to do simulation, if no input is passed time periods will be 
#for next one year, which is 265
timePeriods=forecastDate-Sys.Date()
#creating vector of 1000 simulations with present stock price
simulatedVal=rep(currentStockValue,simulations)
#creating a data frame for the simulations
simulatedDF=data.frame(simulatedVal)
#iterating through random walk, with mean(meanIncrease) and standard deviation (stdIncrease)
for(j in 1:timePeriods)
{
  #creating new vector in every iteration based on the previous simulated values
  #using the same standard deviation and mean we computed from past one year data
  simulatedVal=simulatedVal*replicate(qnorm(runif(1),mean = meanIncrease,sd = stdIncrease),n =simulations)
  #concatenating the computed simulations to the existing data frame
  simulatedDF=cbind(simulatedDF,simulatedVal)
}
#creating a sequence of forecast dates for each day we are predicting
forecastDateNames=seq(from=Sys.Date(),to=forecastDate,by = 1)
#changing column names to the forecasting dates we computed in the above step
colnames(simulatedDF)=forecastDateNames

#function to compute min max at every stage 
#computing difference between the current stock price and forecasted price for every simulation
#if the difference is negative we will return 0, saying we will not make any profit from call option
#if the diiference is more than current value , we are returning it as the increase
minmax=function(earnings)(
  ifelse((earnings-currentStockValue)>0,earnings-currentStockValue,0))
#calling the above function on all the columns(number of time periods using lapply)
simulatedDF=lapply(simulatedDF,minmax)
#converting list output from lapply to a data frame
simulatedDF=as.data.frame(simulatedDF)
#returning the mean income at every point of time, these are our forecasts for the stocks at each point 
#of time
earnings=data.frame(colSums(simulatedDF)/nrow(simulatedDF))
#changing the column name of earnings to earnings
colnames(earnings)=c("earnings")
#giving rownames as the date of forecasts
rownames(earnings)=forecastDateNames
#if the render is true we will return the comlete simulations of the forecasts which is simulatedDF,
##we further use this simulatedDF in a different function
if(render==TRUE)
{
  print("inside render plots")
  results=simulatedDF
  colnames(results)=forecastDateNames
}
else
{
  #this is for the precomputed stocks , where we will only show the forecasts of the mean values
  results=earnings
}
#returning the results
return(results)
}
```
function to compute the forecasts for top 10 stocks
these 10 stocks we will show in the UI as the landing page in shiny application
```{r}
defaultStocks=function()
{
  #stocks to be shown on the landing page
  stocks=list("MSFT","AAPL","GOOGL","IBM","TCS","XOM","HPQ","FB","BAC","JPM")
  #using render as false as this function needs only one mean of the stock predictions
  defaultPredictions=data.frame(
    lapply(stocks, function(stocks)prepareForecast(stockName=stocks,render =FALSE )))
  #converting colnames of the retrieved data frames to respective stock names
  colnames(defaultPredictions)=stocks
  #creating a new column with dates
  defaultPredictions$Date=rownames(defaultPredictions)
  #grouping the predicitons by month and year, to do that computing the month and year from the
  #forecast date
  defaultPredictions=defaultPredictions%>%
    mutate(Month=format(as.Date(Date),"%m"),
           Year=format(as.Date(Date),"%Y"))
  #summarizing with mean monthly stock predictions and sorting it accordingly
  defaultPLot=defaultPredictions%>%dplyr::group_by(Year,Month)%>%
    dplyr::summarise(MSFT=mean(MSFT),AAPL=mean(AAPL),GOOGL=mean(GOOGL),IBM=mean( IBM),
                     TCS=mean(TCS),XOM=mean(XOM),HPQ=mean(HPQ),
                     FB=mean(FB),BAC=mean(BAC),JPM=mean(JPM))%>%
    mutate(time=paste(Year,"-",Month))%>%dplyr::arrange(Year,Month)

#computing and saving the plot to show in landing page of the shiny application
  firstTimeLandingPlot= plot_ly(data=defaultPLot,x=~time,y=~MSFT,name = 'MICROSOFT', type = 'scatter', mode = 'lines')%>%
    add_trace(y = ~AAPL, name = 'AAPLE', mode = 'lines+markers') %>%
    add_trace(y = ~GOOGL, name = 'GOOGLE', mode = 'lines+markers')%>%
    add_trace(y = ~IBM, name = 'IBM', mode = 'lines+markers')%>%
    add_trace(y = ~TCS, name = 'TCS', mode = 'lines+markers')%>%
    add_trace(y = ~XOM, name = 'Exxon Mobil Corp', mode = 'lines+markers')%>%
    add_trace(y = ~HPQ, name = 'HPQ', mode = 'lines+markers')%>%
    add_trace(y = ~FB, name = 'Facebook', mode = 'lines+markers')%>%
    add_trace(y = ~BAC, name = 'Bank of America', mode = 'lines+markers')%>%
    add_trace(y = ~JPM, name = 'CHASE', mode = 'lines+markers')%>%
    layout(xaxis =list(title=" "),
           yaxis =list(title=" "),
           title="Few Stock Performances in the Next One Year")
#returnning the plot to the UI
 return(firstTimeLandingPlot)
}
```

function to render in the UI , shiny application
function which we call from shiny application based on the user input
function flow will be as follows callingRenderingStockForecast--> prepareForecast--> stockForecast
with render as true this function will provide the entire matrix of simulations to shiny
```{r}
callingRenderingStockForecast=function(stockName="Microsoft Corporation",forecastDate="11/26/2017",render=TRUE)
{
  #if the stock has more than one stock , we are taking the first one alone
  symbol=SYMs$Symbol[SYMs$Name==stockName][[1]][1]
  #calling the prepare forecasts function with user values
  forecasts=prepareForecast(stockName =symbol,forecastDate = forecastDate,render = render)
  #returning the complete simulated matrix to the shiny app
return(forecasts)
}
```
